The Power of Unbiased Recursive Partitioning:\\ A Unifying View of CTree, MOB, and GUIDE

### 2019-02-13 pkg/partykit/inst/guideliketest/draft/draft.Rnw
further text revisions and correct path to source code
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2465 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4a9e11b2144c6f6045be53186dd6de2db3dd93eb

%% need no \usepackage{Sweave} ?
%\usepackage{Sweave}
\SweaveOpts{engine = R, concordance = FALSE, eps = FALSE, keep.source = TRUE, eval = TRUE}

@@

<<preliminaries, echo=FALSE, results=hide>>=
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/draft/")
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("partykit")
library("Formula")
library("parallel")
library("lattice")
library("scales")
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/draft/")
source("../guidelike.R")
source("../tests_guideprune.R")

@@

As explained in Section~\ref{sec:introduction} the considered tree algorithms CTree, MOB, and 
GUIDE first select the split variable and then, in a separate step, the split point. 
For the first step of selecting a split variable they all apply statistical tests
following the same general strategy: 
\begin{enumerate}
\item Based on the discrepancy measure an $n \times k$-matrix $D$ is obtained giving information
on the goodness of fit of the model for a data set of $n \in \mathbb{N}$ observations in the 
current node and the parameter $k \in \mathbb{N}$ defined by the algorithm.
\item Using this matrix the algorithm-specific test statistic is evaluated for each possible 
split variable $Z_j$, $j \in \{1,\ldots,m\}$ for $m \in \mathbb{N}$, assessing the dependency
of $D$ on $Z_j$.
\item For the resulting values of the test statistics the corresponding $p$-values are calculated
in order to allow for their comparison on a standardized/unified scale and in that way for
an unbiased variable selection.
%As the variables are usually not all measured at the same scale the resulting
%values of the test statistics can not be compared directly in an unbiased way. Therefore, it is
%necessary to calculate the corresponding $p$-values in order to allow for an unbiased
%variable selection on this standardized/unified scale. 
\item The variable corresponding to the smallest $p$-value is selected as split variable as 
it has shown to be the variable with the strongest influence on the goodness of fit of the model.
\end{enumerate}

@@

%\newpage
<<dgp, echo=FALSE>>=
set.seed(7)

>>>>

%% need no \usepackage{Sweave} 
\SweaveOpts{engine = R, concordance = FALSE, eps = FALSE, keep.source = TRUE, eval = TRUE}

@@

<<preliminaries, echo=FALSE, results=hide>>=
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/draft/")
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

library("partykit")
library("Formula")
library("parallel")
library("lattice")
library("scales")

source("../guidelike.R")
source("../tests_guideprune.R")
source("../ccprune.R")

@@

As explained in Section~\ref{sec:introduction} the considered tree algorithms CTree, MOB, and 
GUIDE first select the split variable and then, in a separate step, the split point. 
For the first step of selecting a split variable they all apply statistical tests
following the same general strategy: 
\begin{enumerate}
\item Based on the discrepancy measure an $n \times k$-matrix $D$ is obtained giving information
on the goodness of fit of the model for a data set of $n \in \mathbb{N}$ observations in the 
current node and the parameter $k \in \mathbb{N}$ defined by the algorithm.
\item Using this matrix the algorithm-specific test statistic is evaluated for each possible 
split variable $Z_j$, $j \in \{1,\ldots,m\}$ for $m \in \mathbb{N}$, assessing the dependency
of $D$ on $Z_j$.
\item For the resulting values of the test statistics the corresponding $p$-values are calculated
in order to allow for their comparison on a standardized/unified scale and in that way for
an unbiased variable selection.
\item The variable corresponding to the smallest $p$-value is selected as split variable as 
it has shown to be the variable with the strongest influence on the goodness of fit of the model.
\end{enumerate}

@@

<<dgp, echo=FALSE>>=
set.seed(7)
